---
params: 
    analysis_title: "DESeq2 differential expression"
title: "`r params$analysis_title`"
author: "VBCF-NGS"
date: "`r format(Sys.time(), '%d %B %Y %H:%M:%S', 'Europe/Vienna', T)`"
output: bookdown::html_document2
---

<style type="text/css">

.figure {
  border: 1px solid black;
  padding: 1em;
  margin-top: 1em;
  margin-bottom: 1em;
}

.figure .figure {
  border: 0px;
  padding: 0px;
  margin: 0px;
}

.book .book-body .page-wrapper .page-inner {
  max-width: 90%;
}

</style>



```{r setup, include=FALSE, cache = FALSE, cache.path=KNITDIR}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE, cache=FALSE, cache.path=KNITDIR)
library("ggplot2")
library("formattable")
library("deseq2analysis")
theme_set(theme_bw())
options(knitr.table.format = "html") 
```

```{r import, include=FALSE, cache=FALSE}
ensembl <- readEnsembl(deseqconfig$ensembltable)
grouping <- readGrouping(deseqconfig$groupingtable) # todo: enable filter grouping in deseqconfig
h2id <- deseqconfig$import$header2id
dds <- readCountsToDeseq2(deseqconfig$countstable, grouping, header2id=get(h2id), remove_genes = NULL, metacols=deseqconfig$import$metacols, skip=deseqconfig$import$skip)
dds.r <- DESeq2::DESeq(dds, betaPrior = TRUE)
comparisons <- readr::read_tsv(deseqconfig$comparisonstable)
deseq.r <- purrrlyr::by_row(comparisons, function(row){ deseqResult(dds.r, row$cond1, row$cond2) }) %>% dplyr::pull(.out)
rlogFull <- DESeq2::rlog(dds.r, blind = TRUE)
```

# DESeq2 analysis: overview


## Output Files and Other Analyses

```{r outputfiles}
   outputFilesTable(deseqconfig, deseq.r) %>%  knitr::kable(caption="output files") %>% kableExtra::kable_styling(bootstrap_options = c("striped", "hover"), latex_options = "striped")
```


## Counts of Deregulated Genes
```{r difffextcounts}

  #color_bar sqishes the numbers
   extractMultiResultsSignif(deseq.r, deseqconfig$deseq2$adjp, lfc <- deseqconfig$deseq2$log2fc)  %>% 
        #dplyr::mutate(dereg=formattable::color_bar("lightgreen")(dereg), up=formattable::color_bar("lightyellow")(up), down=formattable::color_bar("lightorange")(down)) %>%
        knitr::kable(caption="Counts of deregulated genes") %>% kableExtra::kable_styling(bootstrap_options = c("striped", "hover"), latex_options = "striped")

```



```{r totalcounts, fig.cap="Counts per Sample"}
   plotTotalCounts(dds)
```

## Diagnostic Plots Across all Conditions

```{r sizefactors, fig.cap="Size factors used for normalization of the gene counts"}
   plotSizeFactors(dds.r)
```


```{r countspergene, fig.cap="Size factor normalised counts per gene"}
   plotCountsPerGeneBoth(dds.r)
```


```{r allpca, fig.cap="PCA of 500 genes with highest variance across conditions, rLog normalised"}
   DESeq2::plotPCA(rlogFull, intgroup = "group", ntop = 500)
```


```{r deseqpairwise,include=FALSE,message=FALSE,echo=FALSE,warning=FALSE,asis=FALSE}
   
   lfc <- deseqconfig$deseq2$log2fc
   pcut <- deseqconfig$deseq2$adjp 
   

   deseqout <- NULL
   for(deseq.result in deseq.r){

            #comparison was added by me to metadata(deseq.result)
            comparisongroups <- S4Vectors::metadata(deseq.result)$comparison[2:3]
            normcounts <- getCounts(dds.r, comparisongroups, grouping)
            filterThreshold <- getFilterThreshold(deseq.result)

            comp <- toSortedTibble(deseq.result, ensembl, filterThreshold, normcounts)
            comparisonFoldchange <- getComparisonFoldChange(deseq.result)
            tit <- getComparison(deseq.result)
            titfile <- getComparisonString(deseq.result)

            LOG(paste("creating files for: ", tit, titfile, deseqconfig$output$savetables ))
            
            if(deseqconfig$output$savetables){
                saveTables(comp, titfile)
            }

            if(!is.null(deseqconfig$functional)){
                testfunctionalFromConfig(comp, titfile, deseqconfig, tit)
            }

            ADVENTUROUS <- TRUE ## this works better than FALSE where I don't find any output files after the 1. loop.
            if(ADVENTUROUS){
                pairRmd <- knitLabels(titfile, KNITDIR)
            }else{
                pairRmd <- system.file('deseq2_pairwise.Rmd', package="deseq2analysis")
            }

            deseqout <- c(deseqout, knitr::knit_child(pairRmd, options = list(echo = FALSE, warning = FALSE, results = 'hide')))
   }
  
   
  ## cant' find correct environment:
  ## deseqout <- createGroupComparisons(deseqconfig, deseq.r, dds.r, grouping, ensembl)
   
```

# DESeq2 results: pairwise comparisons

To subsequent diagnostic plots are all done with 
an adjusted p-value cutoff of `r pcut` and a minimal
log2 fold change of `r lfc`. This has no influence on the output of the tables which contain all data
sorted by adjusted p-value and log2 fold change. The statistics columns of the tables are given in table \@ref(fig:deseqtabledesc).

```{r deseqtabledesc}
     

```


```{r deseqprint}
   knitr::asis_output(deseqout)
```

# session info
```{r sessioninfo}
   sessioninfo::session_info()
```
