

---
params: 
    functional_title: "Functional Analysis"
title: "`r params$functional_title`"
date: "`r format(Sys.time(), '%B %d, %Y %H:%M:%S', 'America/Chicago', T)`"
output:
  rmarkdown::html_document:
    self_contained: yes
    theme: null
    highlight: null
    mathjax: null

---


```{r setup, echo = FALSE, include=FALSE, cache = FALSE}
library(DOSE)
library(ggplot2)
library("clusterProfiler")
library("deseq2analysis")
options(knitr.table.format = "html") 
theme_set(theme_bw())
outtabledir <- paste(outbasefun,"_tables",sep="")
dir.create(outtabledir)


deseq2parameterfile <- paste(outbasefun,"_deseq2parameters.RDS",sep="")
deseq2parameters <- NULL
if(file.exists(deseq2parameterfile)){
  deseq2parameters <- readRDS(deseq2parameterfile)
}

goup <- readRDS(paste(outbasefun,"_up.RDS",sep=""))
godn <- readRDS(paste(outbasefun,"_dn.RDS",sep=""))
gode <- readRDS(paste(outbasefun,"_de.RDS",sep=""))
gseaL <- readRDS(paste(outbasefun,"_gse.RDS",sep=""))


```

All functional analyses were done with three sets of genes: upregulated genes, downregulated genes and all deregulated genes together. `r goparameterdescription(deseq2parameters)`
The GeneOntology analysis test for overrepresentation of groups of genes within the GO categories. Because the gene-ontology is a directed acyclic graph - picture a branching tree - there is some redundancy in the terms and genes.


# Upregulated Genes

## GO
### GO molecular function
```{r, echo = FALSE}
   emptyTable(goup$mf, datatabGO, "UpGoMf", outtabledir, outbasefun)
```

### GO biological process
```{r, echo = FALSE}
   emptyTable(goup$mf, datatabGO, "UpGoBp", outtabledir, outbasefun)
```

### GO cellular component
```{r, echo = FALSE}
   emptyTable(goup$mf, datatabGO, "UpGoCc", outtabledir, outbasefun)
```

## KEGG
```{r, echo=FALSE}
   emptyTable(goup$kegg, datatabKEGG, "UpKEGG", outtabledir, outbasefun)
```

## Reactome
```{r, echo=FALSE}
   emptyTable(as.data.frame(goup$react), datatabReact, "UpReact", outtabledir, outbasefun)
```
```{r, echo=FALSE}
   plotEmpty(goup$react,  dotplot(goup$react, showCategory=15))
```
```{r, echo=FALSE}
   plotEmpty(goup$react,  emapplot(goup$react))
```


# Downregulated Genes

## GO
### GO molecular function
```{r, echo = FALSE}
   emptyTable(godn$mf, datatabGO, "DnGoMf", outtabledir, outbasefun)
```

### GO biological process
```{r, echo = FALSE}
   emptyTable(godn$bp, datatabGO, "DnGoBb", outtabledir, outbasefun)
```

### GO cellular component
```{r, echo = FALSE}
   emptyTable(godn$cc, datatabGO, "DnGoCc", outtabledir, outbasefun)
```

## KEGG
```{r, echo=FALSE}
   emptyTable(godn$kegg, datatabKEGG, "DnKEGG", outtabledir, outbasefun)
```

## Reactome
```{r, echo=FALSE}
   datatabReact(as.data.frame(godn$react), "DnReact", outtabledir, outbasefun)
```
```{r, echo=FALSE}
   plotEmpty(godn$react,  dotplot(godn$react, showCategory=15))
```
```{r, echo=FALSE}
   plotEmpty(godn$react,  emapplot(godn$react))
```


# Deregulated Genes

## GO
### GO molecular function
```{r, echo = FALSE}
   emptyTable(gode$mf, datatabGO, "DrGoMf", outtabledir, outbasefun)
```

### GO biological process
```{r, echo = FALSE}
   emptyTable(gode$bp, datatabGO, "DrGoBp", outtabledir, outbasefun)
```

### GO cellular component
```{r, echo = FALSE}
   emptyTable(gode$cc, gdatatabGO, "DrGoCc", outtabledir, outbasefun)
```

## KEGG
```{r, echo=FALSE}
   emptyTable(gode$kegg, datatabKEGG, "DrKEGG", outtabledir, outbasefun)
```

## Reactome
```{r, echo=FALSE}
   emptyTable(as.data.frame(gode$react), datatabReact, "DrReact", outtabledir, outbasefun)
```
```{r, echo=FALSE}
   plotEmpty(gode$react,  dotplot(gode$react, showCategory=15))
```
```{r, echo=FALSE}
   plotEmpty(gode$react,  emapplot(gode$react))
```

# Gene Set Enrichment Analysis
For Gene Set Enrichment Analysis (GSEA) no arbitraty cut-offs are used. The whole list, ranked by log-fold change
is compared to gene sets. Multiple Gene Set categories are extracted e.g. the Hallmark gene sets and for each
gene sets the top 10 up and downregulated pathways are shown.

## Hallmark gene sets

Hallmark gene sets  are coherently expressed signatures derived by aggregating many MSigDB gene sets to represent well-defined biological states or processes.

```{r,fig.cap="Hallmark gene sets", echo=FALSE, warnings=FALSE}
   topPathways <- extractGSEAOverviewCollapsedTop(gseaL, 10, "H", gseaL$pvalGo) 
   fgsea::plotGseaTable(gseaL$pathways[topPathways], gseaL$stats, gseaL$gsea, gseaParam = 0.5)
```
```{r, echo=FALSE, warnings=FALSE}
   datatabGSEA(gseaL$gsea, topPathways, "GSEAH", outtabledir, outbasefun)       
```


## Curated gene sets
Curated gene sets from online pathway databases, publications in PubMed, and knowledge of domain experts.

```{r,fig.cap="Curated gene sets", echo=FALSE, warnings=FALSE}
   topPathways <- extractGSEAOverviewCollapsedTop(gseaL, 10, gseaL$pvalGo, "C2")
   fgsea::plotGseaTable(gseaL$pathways[topPathways], gseaL$stats, gseaL$gsea, gseaParam = 0.5)
```

```{r, echo=FALSE, warnings=FALSE}
   datatabGSEA(gseaL$gsea, topPathways, "GSEAC2", outtabledir, outbasefun)
```

## Go gene sets
GO gene sets  consist of genes annotated by the same GO terms.

```{r,fig.cap="GO gene sets", echo=FALSE, warnings=FALSE}
   topPathways <- extractGSEAOverviewCollapsedTop(gseaL, 10, gseaL$pvalGo, "C5")
   fgsea::plotGseaTable(gseaL$pathways[topPathways], gseaL$stats, gseaL$gsea, gseaParam = 0.5)
```

```{r, echo=FALSE, warnings=FALSE}
   datatabGSEA(gseaL$gsea, topPathways, "GSEAC5", outtabledir, outbasefun)
```

## Motif gene sets
Motif gene sets based on conserved cis-regulatory motifs from a comparative analysis of the human, mouse, rat, and dog genomes.

```{r,fig.cap="Motif gene sets", echo=FALSE, warnings=FALSE}
   topPathways <- extractGSEAOverviewCollapsedTop(gseaL, 10, gseaL$pvalGo, "C3")
   fgsea::plotGseaTable(gseaL$pathways[topPathways], gseaL$stats, gseaL$gsea, gseaParam = 0.5)
```

```{r, echo=FALSE, warnings=FALSE}
   datatabGSEA(gseaL$gsea, topPathways, "GSEAC3", outtabledir, outbasefun)
```

# session info
```{r sessioninfo}
   sessioninfo::session_info()
```






